Rising from the Ashes II: Can Bar Formation Lead to Abundance Bimodalities?

### 2024-10-09 paper/ms.tex
Fix for another read through

\newcommand{\rhalf}{\ensuremath{r_{\star,\textrm{half}}}}

\newcommand{\red}[1]{\textcolor{red}{#1}}
% \linespread{1.8}

% \received{March 1, 2021}
% \revised{April 1, 2021}
%\accepted{\today}

\shorttitle{Abundance Bimodality in TNG}
\shortauthors{Beane et al.}

\graphicspath{{./}{fig/}}

@@

\begin{abstract}
    The Milky Way is known to host at least two modes in its present day distribution of Fe and $\alpha$-elements. The exact cause of this bimodality is disputed, but one class of explanations involves the merger between the Milky Way and a relatively massive satellite (Gaia-Sausage-Enceladus) at $z\sim2$. However, reproducing this bimodality in simulations is not straightforward, with conflicting results on the prevalance, morphology, and mechanism behind multimodality. We present a case study of a galaxy in the Illustris TNG50 simulation which undergoes a sequence of starburst, brief quiescence, and then rejuvenation. Following a minor post-processing adjustment -— motivated by recent findings on the star formation efficiency in TNG -- which increases the \alphaFe{} values of older star particles, we demonstrate that this galaxy exhibits a pronounced bimodal distribution in the \alphaFe{}-\FeH{} plane. The high- and low-$\alpha$ sequences are neatly separated in time by the brief quiescent period. The quiescent period in this galaxy is not associated with a merger but by the formation of a bar followed by AGN activity. This galaxy indicates a novel scenario in which the $\alpha$-bimodality is caused by AGN-induced quenching triggered by the formation of the Milky Way's bar. We argue that the post-processing step can be interpreted as the TNG model underproducing star formation in the densest regions at high-$z$.
  \end{abstract}

@@

\section{Introduction}\label{sec:intro}
The stellar surface abundances of most elements retain the composition of their natal gas cloud. Therefore, the present-day distribution of their abundances encodes the chemical history of a galaxy's gas phase. Two classes elements have received particular attention in the Milky Way due to their disparate formation channels: iron-peak (such as Fe) and $\alpha$-elements (produced through the $\alpha$-process, such as O and Mg). Iron-peak elements are produced in both Type~Ia and Type~II supernovae (SNe) whereas $\alpha$-elements are produced predominantly through Type~II SNe. In the Milky Way, there is a well-established bimodality with separate high- and low-$\alpha$ sequences shown in Figure~\ref{fig:fig1} \citep{1996ASPC...92..307G,1998A&A...338..161F,2004AN....325....3F,2006MNRAS.367.1329R,2011A&A...535L..11A,2012A&A...545A..32A,2014A&A...562A..71B,2014ApJ...796...38N,2020MNRAS.493.2952H}.

@@

In \citet{2024arXiv240707985B} (hereafter Paper~I), we proposed an alternate scenario for the formation of the bimodality, driven by a brief ($\sim300\,Myr$) quiescent period in the Galaxy's history, which has two effects. First, the reduced SFR lowers the number of $\alpha$-producing Type~II SNe relative to Type~Ia\footnote{These SNe occur on different timescales (10s of Myr after star formation for Type~II, as compared to 100s of Myr to Gyrs for Type~Ia), so if the SFR drops, the number of Type~II SNe drops faster than Type~Ia.}, causing a sharp decline in the \alphaFe{} ratio in the star-forming gas. Second, the lower SFR results in fewer stars forming in the region between the high- and low-$\alpha$ sequences, reducing the occurance of this transitional population in present-day observations. This mechanism resembles two-phase infall models that incorporate a temporary halt in star formation \citep[][and references therein]{2024arXiv240511025S}, though the quiescent period in Paper~I is much shorter. The combination of these two effects creates the valley between the high- and low-$\alpha$ sequences.

In Paper~I, we used idealized simulations of a galaxy merger that triggered the quiescent period. However, we argued that the merger aspect of that work was not necessary. In this work, we study a subhalo, or galaxy, from the Illustris TNG50 simulation demonstrating this. This subhalo of interest (SoI) exhibits the sequence of events presented in Paper~I after a simple post-processing step which increases the \alphaFe{} of old star particles. This post-processing is motivated by recent work demonstrating that the star formation efficiency (SFE) of the dense gas present at high-$z$ is too low \citep{2024arXiv240909121H}. The SoI undergoes a brief quiescent period which neatly separates a high- and low-$\alpha$ sequence. However, instead of being preceded by a merger, the quenching is preceded by apparent bar-induced AGN activity. Therefore, this work serves as a verification that the scenario in Paper~I is possible in cosmological simulations and does not require a merger.

In Section~\ref{sec:methods} we discuss our selection technique which led to discovering the SoI, the observations we use for comparison, as well as a simple one zone chemical evolution model we use to justify our post-processing step. In Section~\ref{sec:results} we present the main results which we discuss and interpret in Section~\ref{sec:disc}. We conclude in Section~\ref{sec:conc}.

% This decline of \alphaFe{} over time has long been recognized \citep{1979ApJ...229.1046T}. This trend is only generally true, though, with non-monotonic behavior observed in observations and even one-zone chemical models \citep[e.g.][]{2022arXiv220402989C}.

\section{Methods}\label{sec:methods}
\subsection{IllustrisTNG Sample}\label{ssec:tng}
We use Illustris TNG50 \citep{2019MNRAS.490.3196P, 2019MNRAS.490.3234N}, a cosmological simulation of a $\sim50\,\textrm{cMpc}$ box at high resolution ($m_{\textrm{baryon}}\sim8.5\times10^4\,\Msun$). It uses the gravito-magneto-hydrodynamics code \texttt{AREPO} \citep{2010MNRAS.401..791S, 2016MNRAS.455.1134P}, along with the TNG model \citep{2013MNRAS.436.3031V, 2017MNRAS.465.3291W, 2018MNRAS.473.4077P}. This model includes several subgrid processes: a wind generation model, chemical enrichment from SNe and asymptotic giant branch stars, and thermal and kinetic feedback from AGN.

@@

Using the public catalog, we selected a sample of subhalos at $z=1.5$ (snapshot 40) according to the following criteria: (1) the subhalo is central (i.e., the most massive subhalo within its halo), and (2) the subhalo's stellar mass is between $10^{10}$ and $10^{10.5}\,\Msun/\textrm{h}$. There were a total of 168 subhalos that met both criteria. The chosen mass range is broadly consistent with the expected mass of the Milky Way at this redshift \citep{2013ApJ...771L..35V}. We choose to make our selection of galaxies at $z=1.5$ instead of at lower redshift because we wish to capture the \textit{formation} of multimodal structure. Mergers at lower redshift contribute very little to the Milky Way's disk stars \citep[e.g.,][]{2016ARA&A..54..529B}, and would act as a contaminant in our sample.

We examine the abundance distribution in the \MgFe{}-\FeH{} plane for this sample of subhalos by visual inspection. Few subhalos display multimodal structure, and when present, it is much weaker compared to that observed in the Milky Way. We then apply a post-processing step to the \MgFe{} of the stars, adding a value of $0.1\times\left(t_{1.5}-t_{\textrm{form}}\right)$ to each star particle, where $t_{1.5}$ is the age of the universe at $z=1.5$ ($\sim4.3\,\Gyr$), and $t_{\textrm{form}}$ is the formation time of the star particle. After this adjustment, which we justify in Section~\ref{ssec:sfe} based on TNG underpredicting the SFE of dense gas, we observe that the abundance structure becomes significantly more pronounced.

We show the abundance distributions (replicating Figure~\ref{fig:fig1}) of 16 random subhalos from our catalog, and the subhalo we selected, in Appendix~\ref{app:rand_fig1}. Of these subhalos, some structure is present, but none have a strong bimodality as seen in the Milky Way. After the $\alpha$-enhancement, a few have strongly bimodal features (subject to interpretation): 172175, 178140, 193025.

\begin{table}[]
  \centering
  \begin{tabular}{ll}
  \multicolumn{2}{c}{dark matter}                                                                  \\ \hline
  \multicolumn{1}{l|}{$M_{200}$}                                      & $4.97\times10^{12}\,\Msun$ \\
  \multicolumn{1}{l|}{$R_{200}$}                                      & $532\,\kpc$                \\
                                                                      &                            \\
  \multicolumn{2}{c}{baryons}                                                                      \\ \hline
  \multicolumn{1}{l|}{$\rhalf$}                                       & $2\,\kpc$                  \\
  \multicolumn{1}{l|}{$M_{\textrm{star}}(r<4\rhalf)$}                 & $7.6\times10^{12}\,\Msun$  \\
  \multicolumn{1}{l|}{$M_{\textrm{gas}}(r<4\rhalf)$}                  & $2.6\times10^{11}\,\Msun$  \\
  \multicolumn{1}{l|}{$\textrm{SFR}$}                                 & $0.56\,\Msunyr$            \\
  \multicolumn{1}{l|}{$M_{\textrm{BH}}$}                              & $2.9\times10^{8}\,\Msun$   \\
  \multicolumn{1}{l|}{$\dot{M}_{\textrm{BH}}/\dot{M}_{\textrm{Edd}}$} & $4.5\times10^{-5}$        
  \end{tabular}
  \caption{Summary statistics of our subhalo of interest at $z=0$. $M_{200}$ and $R_{200}$ are relative to the mean density of the universe, the SFR is for all gas bound to the subhalo, and the BH refers to the central BH.}
  \label{tab:summ}
\end{table}

@@

\subsection{Observations}\label{ssec:obs}
We make use of two observational data sets. First, we use the ASPCAP DR17 catalog of stellar abundances \citep[][J.A.~Holtzman et al., in preparation]{2016AJ....151..144G}. We make the same selection cuts as in Section~2.4 of Paper I. These criteria are meant to select giants with high quality abundance measurements as well as restricting the sample to only stars in the disk with angular momenta similar to the Sun's. This results in a sample of 54,777 stars. We use Fe to track total metallicity and Mg alone as an $\alpha$-element.

We then further considered a dataset of stellar ages from the APOKASC2 catalog \citep{2018ApJS..239...32P}. This catalog uses a combination of APOGEE spectroscopic parameters and \textit{Kepler} time series photometry to compute astroseismic ages. Using only stars with $25\%$ age uncertainties (taken as the maximum of the upper and lower uncertainty), we cross-match this catalog to our larger sample from ASPCAP which results in a sample of 1201 stars.

\subsection{One-Zone Chemical Evolution Model}\label{ssec:onezone_met}
One-zone galactic chemical evolution (GCE) models describe enrichment in an idealized gas reservoir in which newly synthesized metals mix instantaneously (see e.g. the reviews by \citealt{Tinsley1980} and \citealt{Matteucci2021}). Our parameter choices in this paper are based on the models in \citet{2022arXiv220402989C}. While they were interested in bursts of star formation at the transition between the stellar halo and thick disk formation epochs, we focus on smooth evolutionary histories in this paper. We integrate these models numerically using the publicly available {\tt Versatile Integrator for Chemical Evolution} ({\tt VICE}; \citealt{2020MNRAS.498.1364J}).

The quantity that is of particular importance to our results is the star formation efficiency (SFE) timescale $\tau_\star$:
\begin{equation}
\tau_\star \equiv \frac{M_{\textrm{gas}}}{\dot{M}_{\textrm{star}}},
\end{equation}
where $M_{\textrm{gas}}$ and $\dot{M}_{\textrm{star}}$ are the ISM mass and star formation rate, respectively. This timescale can be equivalently expressed as the ratio of the corresponding surface densities $\Sigma_{\textrm{gas}}$ and $\dot{\Sigma}_{\textrm{star}}$. $\tau_\star$ is the inverse of the SFE itself, describing the rate of star formation per unit mass in the ISM. In principle, $\tau_\star$ should vary with $M_{\textrm{gas}}$ because the observed relation between $\dot{\Sigma}_{\textrm{star}}$ and $\Sigma_{\textrm{gas}}$ is non-linear; classically, $\dot{\Sigma}_{\textrm{star}} \propto \Sigma_{\textrm{gas}}^N$ where $N \approx 1.5$ (e.g. \citealt{Schmidt1959, Kennicutt1998}). For simplicity, we hold $\tau_\star$ constant in this paper, which corresponds to a linear relation. In section 3.5 below, we present models using $\tau_\star = 0.5$, $1$, and $5$ Gyr.

Following \citet{2022arXiv220402989C}, our models assume that there is initially no gas present in the galaxy (i.e. $M_{\textrm{gas}} = 0$ at $t = 0$). Zero metallicity gas accretes at a constant rate of $\dot{M}_\text{in} = 5\,\Msunyr$. The efficiency with which feedback sweeps up ISM material and ejects it to the CGM in an outflow is described by the mass loading factor $\eta \equiv \dot{M}_\text{out} / \dot{M}_\star$. We use $\eta = 2$ throughout this paper (i.e. for every solar mass of stars formed, $2\,\Msun$ of ISM gas is lost to an outflow). These parameter choices lead to an SFH that initially rises from zero and approaches a constant value of $\sim1.9\,\Msunyr$. With these choices, the predicted abundance evolution is insensitive to the overall normalization of the accretion and star formation histories, since a larger accretion rate forms more stars and metals but simultaneously introduces more H into the ISM.

We use the solar abundances of Fe and Mg from \citet{Magg2022} with an additional 0.04 dex to account for the effects of diffusion and gravitational settling (\citealt{Turcotte1998}; i.e. $\{Z_{\text{Fe},\odot}, Z_{\text{Mg},\odot}\} = \{13.7, 6.71\} \times 10^{-4}$). We adopt SN yields of Fe and Mg based on \citeauthor{Weinberg2023}'s \citeyearpar{Weinberg2023} recommendations for Type~II SNe:
\begin{itemize}

@@

\end{itemize}
where the subscripts and superscripts denote the element and SN type, respectively. We select this value of $y_\text{Fe}^\text{Ia}$ as it corresponds to $\sim30\%$ ($\sim70\%$) of Fe arising from Type~II (Ia) SNe at solar \FeH{}. These yields are population-averaged, meaning that they quantify the amount of metal production per unit mass of star formation under the given enrichment channel (e.g. a hypothetical $1000\,\Msun$ stellar population would produce a total of $0.652\,\Msun$ of Mg through Type~II SNe). {\tt VICE} models Type~II SNe as exploding instantaneously following an episode of star formation. For Type~Ia SNe, we assume that metals are produced according to the empirically motivated $t^{-1.1}$ power-law based volumetric SN rates as a function of redshift and the cosmic SFH \citep[e.g.][]{Maoz2012}.

@@

The main result of our paper is given in Figure~\ref{fig:fig1}. Here, we compare the abundance plane in the Milky Way (left column) to that of our subhalo of interest (middle and right columns). The upper panels show the 2D distribution in the space of \MgFe{}-\FeH{}. We have applied the standard \texttt{scipy} implementation of a gaussian kernel density estimator to a Cartesian grid of points. For each panel, we normalize so that the integral of the distribution is unity. Colors are plotted in a log scale ranging from $0.08$ to $15\,\dex^{-2}$. Contour lines are plotted at $0.1$, $1.5$, and $10\,\dex^{-2}$.

The colored vertical regions are indicated at $\FeH=-0.75$, $-0.5$, $-0.25$, and $0\,\dex$ in the Milky Way, and at bins $0.25\,\dex$ higher in the simulations. The lower panels show 1D histograms of \MgFe{} in bins centered on these values. The bins for the Milky Way/simulations have width $0.2$/$0.05\,\dex$. The rationale for the higher plotted \FeH{} in the simulations comes just from the empirical location of the bimodalities. The Milky Way shows a clear bimodal population, with a high-$\alpha$ sequence most clearly distinct from the low-$\alpha$ sequence at low metallicity. The two sequences merge around solar metallicity.

Our SoI, on the other hand, does not show a clearly bimodal structure in the fiducial simulation (middle column). There is some structure in the $\FeH=-0.75$ bin. The right panel of Figure~\ref{fig:fig1} shows the same distribution as in the middle panel, but with a post-processed declination in \MgFe{} described in Section~\ref{ssec:tng}. Star particles formed before $z=1.5$ are given an additive offset of $0.1\times\left(t_{1.5}-t_{\textrm{form}}\right)$, where $t_{1.5}$ is the age of the universe at $z=1.5$ and $t_{\textrm{form}}$ is the formation time of the star particle. A multimodal structure emerges with three clear modes at $\MgFe\sim0.8$, $0.5$, and $0.2\,\dex$. The 1D histograms show that the modes are well-separated, and that the troughs between the modes nearly vanish.

@@

In the lower row, the left panel shows data from the Milky Way, while the middle and right panels correspond to the fiducial and $\alpha$-enhanced SoI, respectively. For the simulations, we introduce Gaussian errors of $15\%$ in age and $0.01,\dex$ in \MgFe{}. These error values align with the observational uncertainties from the APOKASC2 and APOGEE datasets, which the Milky Way panel (lower left) is based on (see Appendix~\ref{app:obs_err} for detailed plots of observational errors). While the error in \MgFe{} is relatively minor, the age error ($1.5\,\Gyr$ at $10\,\Gyr$) blurs the distribution, particularly around the dashed line that marks the transition between sequences. Despite this, the $\alpha$-enhanced SoI still shows two distinct populations, although their ages now overlap more significantly.

@@

\subsection{Evolutionary History}\label{ssec:evol}
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{fig/392276_SFH_AGN_M200.pdf}
  \caption{\textbf{The evolutionary history of our subhalo of interest.} The left column shows the SFH, BH accretion rate, and virial mass ($M_{200}$) over the entire time span, while the right column zooms in on the period from $t\sim2\,\Gyr$ to $t\sim5\,\Gyr$. The vertical dashed line in each panel marks the transition at $t\sim3.2\,\Gyr$, corresponding to the separation between the high- and low-$\alpha$ sequences (as shown in Figure~\ref{fig:alpha}). The middle panel shows the BH accretion rate as a fraction of the Eddington limit, and the lower panel shows the virial mass, representing the mass enclosed within a radius where the density is $200\times$ the mean cosmic density.}
  \label{fig:history}
\end{figure}

>>>>

\newcommand{\rhalf}{\ensuremath{r_{\star,\textrm{half}}}}
\newcommand{\SFR}{\ensuremath{\textrm{SFR}}}

\newcommand{\red}[1]{\textcolor{red}{#1}}

@@

\begin{abstract}
    The Milky Way hosts at least two modes in its present day distribution of Fe and $\alpha$-elements. The exact cause of this bimodality is disputed, but one class of explanations involves the merger between the Milky Way and a relatively massive satellite (Gaia-Sausage-Enceladus) at $z\sim2$. However, reproducing this bimodality in simulations is not straightforward, with conflicting results on the prevalance, morphology, and mechanism behind multimodality. We present a case study of a galaxy in the Illustris TNG50 simulation which undergoes a sequence of starburst, brief quiescence, and then rejuvenation. Following a minor post-processing adjustment -— motivated by recent findings on the star formation efficiency in TNG -- which increases the \alphaFe{} values of older star particles, we demonstrate that this galaxy exhibits a pronounced bimodal distribution in the \alphaFe{}-\FeH{} plane. The high- and low-$\alpha$ sequences are neatly separated in time by the brief quiescent period. The quiescent period in this galaxy is not associated with a merger but by the formation of a bar followed by AGN activity. This galaxy indicates a novel scenario in which the $\alpha$-bimodality in the Milky Way is caused by AGN-induced quenching triggered by the formation of its bar. We predict that abundance bimodalities should be more common in barred as opposed to unbarred galaxies.
  \end{abstract}

@@

\section{Introduction}\label{sec:intro}
The stellar surface abundances of most elements retain the composition of their natal gas cloud. Therefore, the present-day distribution of their abundances encodes the chemical history of a galaxy's gas phase. Two classes of elements have received particular attention in the Milky Way due to their disparate formation channels: iron-peak (such as Fe) and $\alpha$-elements (produced through the $\alpha$-process, such as O and Mg). Iron-peak elements are produced in both Type~Ia and Type~II supernovae (SNe) whereas $\alpha$-elements are produced predominantly through Type~II SNe. In the Milky Way, there is a well-established bimodality with separate high- and low-$\alpha$ sequences shown in Figure~\ref{fig:fig1} \citep{1996ASPC...92..307G,1998A&A...338..161F,2004AN....325....3F,2006MNRAS.367.1329R,2011A&A...535L..11A,2012A&A...545A..32A,2014A&A...562A..71B,2014ApJ...796...38N,2020MNRAS.493.2952H}.

@@

In \citet{2024arXiv240707985B} (hereafter Paper~I), we proposed an alternate scenario for the formation of the bimodality, driven by a brief ($\sim300\,\Myr$) quiescent period in the Galaxy's history, which has two effects. First, the reduced SFR lowers the number of $\alpha$-producing Type~II SNe relative to Type~Ia\footnote{These SNe occur on different timescales (10s of Myr after star formation for Type~II, as compared to 100s of Myr to Gyrs for Type~Ia), so if the SFR drops, the number of Type~II SNe drops faster than Type~Ia.}, causing a sharp decline in the \alphaFe{} ratio in the star-forming gas. Second, the lower SFR results in fewer stars forming in the region between the high- and low-$\alpha$ sequences, reducing the occurance of this transitional population in present-day observations. This mechanism resembles two-phase infall models that incorporate a temporary halt in star formation \citep[][and references therein]{2024arXiv240511025S}, though the quiescent period in Paper~I is much shorter. The combination of these two effects creates the valley between the high- and low-$\alpha$ sequences.

In Paper~I, we used idealized simulations of a galaxy merger that triggered the quiescent period. However, in that work we argued the merger was unnecessary. In this work, we study a subhalo, or galaxy, from the Illustris TNG50 simulation demonstrating this. This subhalo of interest (SoI) exhibits the sequence of events presented in Paper~I after a simple post-processing step which increases the \alphaFe{} of old star particles. This post-processing is motivated by recent work demonstrating that the star formation efficiency (SFE) of the dense gas present at high-$z$ is too low in the TNG model \citep{2024arXiv240909121H}. The SoI undergoes a brief quiescent period which neatly separates a high- and low-$\alpha$ sequence. However, instead of being preceded by a merger, the quenching is preceded by apparent bar-induced AGN activity. Therefore, this work serves as a verification that the scenario in Paper~I is possible in cosmological simulations and does not require a merger.

In Section~\ref{sec:methods}, we discuss our selection technique which led to discovering the SoI, the observations we use for comparison, as well as a simple one zone chemical evolution model we use to justify our post-processing step. In Section~\ref{sec:results}, we present the main results which we discuss and interpret in Section~\ref{sec:disc}. We conclude in Section~\ref{sec:conc}.

\section{Methods}\label{sec:methods}
\subsection{IllustrisTNG Sample}\label{ssec:tng}
We use Illustris TNG50 \citep{2019MNRAS.490.3196P, 2019MNRAS.490.3234N}, a cosmological simulation of a $\sim50\,\textrm{cMpc}$ box at high resolution ($m_{\textrm{baryon}}\sim8.5\times10^4\,\Msun$). It uses the gravito-magneto-hydrodynamics code \texttt{AREPO} \citep{2010MNRAS.401..791S, 2016MNRAS.455.1134P}, along with the TNG model \citep{2013MNRAS.436.3031V, 2017MNRAS.465.3291W, 2018MNRAS.473.4077P}. This model includes several subgrid processes, including a wind generation model, chemical enrichment from SNe and asymptotic giant branch stars, and thermal and kinetic feedback from AGN.

@@

Using the public catalog, we select a sample of subhalos at $z=1.5$ (snapshot 40) according to the following criteria: (1) the subhalo is central (i.e., the most massive subhalo within its halo), and (2) the subhalo's stellar mass is between $10^{10}$ and $10^{10.5}\,\Msun/\textrm{h}$. There were a total of 168 subhalos that met both criteria. The chosen mass range is broadly consistent with the expected mass of the Milky Way from abundance matching at this redshift \citep{2013ApJ...771L..35V}. We choose to make our selection of galaxies at $z=1.5$ instead of at lower redshift because we wish to capture the \textit{formation} of multimodal structure. Mergers at lower redshift contribute very little to the Milky Way's disk stars \citep[e.g.,][]{2016ARA&A..54..529B}, and would act as a contaminant in our sample.

We examine the abundance distribution in the \MgFe{}-\FeH{} plane for this sample of subhalos by visual inspection. Few subhalos display multimodal structure and, when present, is much weaker compared to that observed in the Milky Way. We then apply a post-processing step to the \MgFe{} of the stars, adding a value of $0.1\times\left(t_{1.5}-t_{\textrm{form}}\right)$ to each star particle, where $t_{1.5}$ is the age of the universe at $z=1.5$ ($\sim4.3\,\Gyr$), and $t_{\textrm{form}}$ is the formation time of the star particle. After this adjustment, which we justify in Section~\ref{ssec:sfe} based on TNG underpredicting the SFE of dense gas, we observe that the abundance structure becomes significantly more pronounced.

We show the abundance distributions (replicating Figure~\ref{fig:fig1}) of 16 random subhalos from our catalog, and the subhalo we selected, in Appendix~\ref{app:rand_fig1}. Of these subhalos, some structure is present, but none have a strong bimodality as seen in the Milky Way. After the $\alpha$-enhancement, a few have strongly bimodal features: 172175, 178140, 193025.

\begin{table}[htb]
  \centering
  \begin{tabular}{ll}
  \multicolumn{2}{c}{dark matter}                                                                  \\ \hline
  \multicolumn{1}{l|}{$M_{200}$}                                      & $4.97\times10^{12}\,\Msun$ \\
  \multicolumn{1}{l|}{$R_{200}$}                                      & $532\,\kpc$                \\
                                                                      &                            \\
  \multicolumn{2}{c}{baryons}                                                                      \\ \hline
  \multicolumn{1}{l|}{$\rhalf$}                                       & $2\,\kpc$                  \\
  \multicolumn{1}{l|}{$M_{\textrm{star}}(r<4\rhalf)$}                 & $7.6\times10^{12}\,\Msun$  \\
  \multicolumn{1}{l|}{$M_{\textrm{gas}}(r<4\rhalf)$}                  & $2.6\times10^{11}\,\Msun$  \\
  \multicolumn{1}{l|}{$\textrm{SFR}$}                                 & $0.56\,\Msunyr$            \\
  \multicolumn{1}{l|}{$M_{\textrm{BH}}$}                              & $2.9\times10^{8}\,\Msun$   \\
  \multicolumn{1}{l|}{$\dot{M}_{\textrm{BH}}/\dot{M}_{\textrm{Edd}}$} & $4.5\times10^{-5}$        
  \end{tabular}
  \caption{Summary statistics of our subhalo of interest at $z=0$ (392276). $M_{200}$ and $R_{200}$ are relative to the mean density of the universe, the SFR is for all gas bound to the subhalo, and the BH refers to the central BH.}
  \label{tab:summ}
\end{table}

@@

\subsection{Observations}\label{ssec:obs}
We make use of two observational data sets. First, we use the ASPCAP DR17 catalog of stellar abundances \citep[][J.A.~Holtzman et al., in preparation]{2016AJ....151..144G}. We make the same selection cuts as in Section~2.4 of Paper I. These criteria select giants with high quality abundance measurements and angular momenta similar to the Sun's. This results in a sample of 54,777 stars. We use Fe to track total metallicity and Mg alone as an $\alpha$-element.

We then further considered a dataset of stellar ages from the APOKASC-3 catalog \citep{2024arXiv241000102P}. This catalog uses a combination of APOGEE spectroscopic parameters and \textit{Kepler} time series photometry to compute astroseismic ages. Using only stars with $25\%$ age uncertainties (taken as the maximum of the upper and lower uncertainty), we cross-match this catalog to our larger sample from ASPCAP which results in a sample of 1201 stars.

\subsection{One-Zone Chemical Evolution Model}\label{ssec:onezone_met}
One-zone galactic chemical evolution (GCE) models describe enrichment in an idealized gas reservoir in which newly synthesized metals mix instantaneously (see, e.g., the reviews by \citealt{Tinsley1980} and \citealt{Matteucci2021}). Our parameter choices in this paper are based on the models in \citet{2022arXiv220402989C}. While they were interested in bursts of star formation at the transition between the stellar halo and thick disk formation epochs, we focus on smooth evolutionary histories in this paper. We integrate these models numerically using the publicly available {\tt Versatile Integrator for Chemical Evolution} ({\tt VICE}; \citealt{2020MNRAS.498.1364J}).

The quantity of particular importance to our results is the star formation efficiency (SFE) timescale $\tau_\star$:
\begin{equation}
\tau_\star \equiv \frac{M_{\textrm{gas}}}{\SFR},
\end{equation}
where $M_{\textrm{gas}}$ is the mass of the gas reservoir. This timescale can be equivalently expressed as the ratio of the corresponding surface densities $\Sigma_{\textrm{gas}}$ and $\dot{\Sigma}_{\textrm{star}}$. $\tau_\star$ is the inverse of the SFE itself, describing the rate of star formation per unit gas mass. In principle, $\tau_\star$ should vary with $M_{\textrm{gas}}$ because the observed relation between $\dot{\Sigma}_{\textrm{star}}$ and $\Sigma_{\textrm{gas}}$ is non-linear; classically, $\dot{\Sigma}_{\textrm{star}} \propto \Sigma_{\textrm{gas}}^N$ where $N \approx 1.5$ (e.g. \citealt{Schmidt1959, Kennicutt1998}). For simplicity, we hold $\tau_\star$ constant in this paper, which corresponds to a linear relation. In Section~\ref{ssec:onezone} below, we present models using $\tau_\star = 0.5$, $1$, and $5\,\Gyr$.

Following \citet{2022arXiv220402989C}, our models assume that there is initially no gas present in the galaxy (i.e. $M_{\textrm{gas}} = 0$ at $t = 0$). Zero metallicity gas accretes at a constant rate of $\dot{M}_\text{in} = 5\,\Msunyr$. The efficiency with which feedback sweeps up interstellar material and ejects it to the circumgalactic medium in an outflow is described by the mass loading factor $\eta \equiv \dot{M}_\text{out} / \SFR$. We use $\eta = 2$ throughout this paper (i.e. for every solar mass of stars formed, $2\,\Msun$ of interstellar gas is lost to an outflow). These parameter choices lead to an SFH that initially rises from zero and approaches a constant value of $\sim1.9\,\Msunyr$. With these choices, the predicted abundance evolution is insensitive to the overall normalization of the accretion and star formation histories, since a larger accretion rate forms more stars and metals but simultaneously introduces more H into the interstellar medium.

We use the solar abundances of Fe and Mg from \citet{Magg2022} with an additional $0.04\,\dex$ to account for the effects of diffusion and gravitational settling (\citealt{Turcotte1998}; i.e. $\{Z_{\text{Fe},\odot}, Z_{\text{Mg},\odot}\} = \{13.7, 6.71\} \times 10^{-4}$). We adopt SN yields of Fe and Mg based on \citeauthor{Weinberg2023}'s \citeyearpar{Weinberg2023} recommendations for Type~II SNe:
\begin{itemize}

@@

\end{itemize}
where the subscripts and superscripts denote the element and SN type, respectively. We select this value of $y_\text{Fe}^\text{Ia}$ as it corresponds to $\sim30\%$ ($\sim70\%$) of Fe arising from Type~II (Ia) SNe at solar \FeH{}. These yields are population-averaged, meaning that they quantify the amount of metal production per unit mass of star formation under the given enrichment channel (e.g. a hypothetical $10^3\,\Msun$ stellar population would produce a total of $0.652\,\Msun$ of Mg through Type~II SNe). {\tt VICE} models Type~II SNe as exploding instantaneously following an episode of star formation. For Type~Ia SNe, we assume that metals are produced according to the empirically motivated $t^{-1.1}$ power-law based volumetric SN rates as a function of redshift and the cosmic SFH \citep[e.g.][]{Maoz2012}.

@@

The main result of our paper is given in Figure~\ref{fig:fig1}. Here, we compare the abundance plane in the Milky Way (left column) to that of our subhalo of interest (middle and right columns). The upper panels show the 2D distribution in the space of \MgFe{}-\FeH{}. We have applied the standard \texttt{scipy} implementation of a gaussian kernel density estimator to a Cartesian grid of points. For each panel, we normalize so that the integral of the distribution is unity. Colors are plotted in a log scale ranging from $0.08$ to $15\,\dex^{-2}$. Dashed contour lines are plotted at $0.1$, $1.5$, and $10\,\dex^{-2}$.

The colored vertical regions are indicated at $\FeH=-0.75$, $-0.5$, $-0.25$, and $0\,\dex$ in the Milky Way, and at bins $0.25\,\dex$ higher in the simulations. The lower panels show 1D histograms of \MgFe{} in bins centered on these values. The bins have width $0.2$/$0.05\,\dex$, which is reflected in the width of the colored, shaded regions in the upper panels. The rationale for the higher plotted \FeH{} in the simulations reflects the empirical location of the bimodalities. The Milky Way shows a clear bimodal population, with a high-$\alpha$ sequence most clearly distinct from the low-$\alpha$ sequence at low metallicity. The two sequences merge around solar metallicity.

The SoI, on the other hand, does not show a clearly bimodal structure in the fiducial simulation (middle column). There is some structure in the $\FeH=-0.75$ bin. The right panel of Figure~\ref{fig:fig1} shows the same distribution as in the middle panel, but with a modification to increase \MgFe{} values of older stars formed before $z=1.5$ (see Sections~\ref{ssec:tng} and \ref{ssec:onezone}). A multimodal structure emerges with three clear modes at $\MgFe\sim0.8$, $0.5$, and $0.2\,\dex$. The 1D histograms show that the modes are well-separated, and that the troughs between the modes nearly vanish.

@@

In the lower row, the left panel shows data from the Milky Way, while the middle and right panels correspond to the fiducial and $\alpha$-enhanced SoI, respectively. For the simulations, we introduce Gaussian errors of $15\%$ in age and $0.01,\dex$ in \MgFe{}. These error values align with the observational uncertainties from the APOKASC2 and APOGEE datasets, which the Milky Way panel (lower left) is based on (see Appendix~\ref{app:obs_err} for detailed plots of observational errors). While the error in \MgFe{} is insignificant, the age error ($1.5\,\Gyr$ at $10\,\Gyr$) blurs the distribution, particularly around the dashed line that marks the transition between sequences. Despite this, the $\alpha$-enhanced SoI still shows two distinct populations, although their ages now overlap more significantly.

@@

\subsection{Evolutionary History}\label{ssec:evol}
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{fig/392276_SFH_AGN_M200.pdf}
  \caption{\textbf{The evolutionary history of our subhalo of interest.} The left column shows the SFH, BH accretion rate, and virial mass ($M_{200}$) over the entire time span, while the right column zooms in on the period from $t=2\,\Gyr$ to $t=5\,\Gyr$. The vertical dashed line in each panel marks the transition at $t\sim3.2\,\Gyr$, corresponding to the separation between the high- and low-$\alpha$ sequences (as shown in Figure~\ref{fig:alpha}). The upper panel shows the SFH, the middle panel shows the BH accretion rate as a fraction of the Eddington limit, and the lower panel shows the virial mass, representing the mass enclosed within a radius where the density is $200\times$ the mean cosmic density.}
  \label{fig:history}
\end{figure}

To understand the key events driving the behavior around the dashed line in Figure~\ref{fig:alpha}, we examine the evolution of several properties of the SoI in Figure~\ref{fig:history}. The vertical dashed line in each panel marks the transition at $t\sim3.2\,\Gyr$ from the high- to low-$\alpha$ sequences, as in Figure~\ref{fig:alpha}. The upper panel in the left column shows the the SFR (computed for all gas cells in the SoI). There are two peaks at $t\sim2.5\,\Gyr$ and $t\sim4.5\,\Gyr$ with maximum values of $50\,\Msun/\textrm{yr}$ and $30\,\Msun/\textrm{yr}$, respectively. Around the high- to low-$\alpha$ transition, there is a dip in the SFR, which drops by an order of magnitude to about $3\,\Msun/\textrm{yr}$. The right panel zooms in on the period between $t=2\,\Gyr$ and $5\,\Gyr$, where we observe a sharp recovery in the SFR following the quiescent phase. In the span of a single snapshot (roughly $150\,\Myr$), the SFR increases from about $3\,\Msun/\textrm{yr}$ to $10\,\Msun/\textrm{yr}$.
